---
title: "Project 3: p3package Tutorial"
author: "Joia Zhang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{p3package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The p3package package contains functions for statistical prediction and statistical inference. Included are one sample t-test, linear model, K-nearest neighbors, and random forest functions. Such functions may be useful for conducting hypothesis testing, modeling, and prediction.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

You can install the p3package package using the following line:

```r{eval = FALSE}
# install.packages("devtools")
devtools::install_github("joiazhang/p3package", build_vignette = TRUE, build_opts = c())
library(p3package)
```

## Tutorials

### `my_t.test` Tutorial
A test of the hypothesis:
\begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}
```{r}
my_gapminder$lifeExp
```




